\chapter{Results and Discussions}\label{ch:4}

This chapter presents the analysis and interpretation of the data gathered from the respondents regarding the developed web-based learning tool, \textit{StatiCalcs}. The findings focus on the respondents’ perceptions in terms of usability, accessibility, and satisfaction, as well as their preference for different learning resources in studying Statics of Rigid Bodies. Frequency distribution, weighted mean, and independent samples t-test were utilized to analyze the gathered data.

%==========Students’ Perception on Usability=========================

\section{Level of Perception of Students Regarding StatiCalcs in Terms of Usability}

\begin{table}[H]
\centering
\caption{Weighted Mean of Students’ Perception on Usability}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{cp{6.5cm}cc}
\hline
\textbf{No.} & \textbf{Statement} & \textbf{Weighted Mean} & \textbf{Interpretation}\\
\hline
1 & The StatiCalcs website is easy to navigate. & 4.69 & Strongly Agree\\
2 & I can complete calculations quickly using the website. & 4.57 & Strongly Agree\\
3 & The website provides clear feedback after performing calculations. & 4.66 & Strongly Agree\\
4 & I can easily correct mistakes when entering inputs in the calculators. & 4.70 & Strongly Agree\\
5 & Overall, StatiCalcs is user-friendly. & 4.71 & Strongly Agree\\
\multicolumn{2}{l}{\textbf{Overall Weighted Mean}} & \textbf{4.66} & \textbf{Strongly Agree}\\
\hline
\end{tabular}
\end{table}

The findings reveal that the student respondents strongly agreed that StatiCalcs is usable and user-friendly, with an overall weighted mean of 4.66 interpreted as ``Strongly Agree.'' The statement ``Overall, StatiCalcs is user-friendly'' obtained the highest weighted mean of 4.71, indicating that the respondents positively evaluated the ease of use and navigation of the platform. Meanwhile, the statement ``I can complete calculations quickly using the website'' obtained the lowest weighted mean of 4.57, although it still remained within the ``Strongly Agree'' interpretation range. The results suggest that the developed web-based learning tool effectively supported efficient interaction and problem-solving among students.

These findings align with Nielsen’s Usability Heuristics, which emphasize that systems should be easy to navigate, efficient to use, and capable of helping users accomplish tasks effectively \autocite{nielsen_10_1994}. Similarly, \cite{hassan_usability_2022} stated that digital learning tools become more effective when interfaces are simple, visually clear, and easy to navigate, allowing learners to focus on learning tasks rather than platform complexity.
%==========Students’ Perception on Accessibility=========================

\section{Level of Perception of Students Regarding StatiCalcs in Terms of Accessibility}

\begin{table}[H]
\centering
\caption{Weighted Mean of Students’ Perception on Accessibility}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{cp{6.5cm}cc}
\hline
\textbf{No.} & \textbf{Statement} & \textbf{Weighted Mean} & \textbf{Interpretation}\\
\hline
1 & The text and visual elements of StatiCalcs are easy to read. & 4.66 & Strongly Agree\\
2 & The website works properly on different devices such as laptops and mobile phones. & 4.66 & Strongly Agree\\
3 & The layout makes important information easy to locate. & 4.69 & Strongly Agree\\
4 & The explanations and results provided by the website are easy to understand. & 4.61 & Strongly Agree\\
5 & Overall, StatiCalcs is accessible under different conditions. & 4.63 & Strongly Agree\\
\multicolumn{2}{l}{\textbf{Overall Weighted Mean}} & \textbf{4.65} & \textbf{Strongly Agree}\\
\hline
\end{tabular}
\end{table}

The results indicate that the student respondents strongly agreed that StatiCalcs is accessible and functional across different devices and learning conditions, with an overall weighted mean of 4.65 interpreted as ``Strongly Agree.'' The statement ``The layout makes important information easy to locate'' obtained the highest weighted mean of 4.69, suggesting that the organization and presentation of information supported ease of access and navigation. Meanwhile, the statement ``The explanations and results provided by the website are easy to understand'' obtained the lowest weighted mean of 4.61, although all indicators remained within the ``Strongly Agree'' interpretation range.

The findings support the Web Content Accessibility Guidelines (WCAG 2.1) of the \cite{world_wide_web_consortium_w3c_web_nodate}, which emphasize that digital platforms should be perceivable, operable, understandable, and accessible across various user conditions and devices. Furthermore, \cite{hassan_usability_2022} highlighted that accessible and user-friendly digital learning tools reduce cognitive load and improve learner engagement by making information easier to locate and understand.

%==========Students’ Perception on Satisfaction=========================

\section{Level of Perception of Students Regarding StatiCalcs in Terms of Satisfaction}

\begin{table}[H]
\centering
\caption{Weighted Mean of Students’ Perception on Satisfaction}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{cp{6.5cm}cc}
\hline
\textbf{No.} & \textbf{Statement} & \textbf{Weighted Mean} & \textbf{Interpretation}\\
\hline
1 & I am satisfied with my overall experience using StatiCalcs. & 4.69 & Strongly Agree\\
2 & StatiCalcs meets my expectations as a supplementary learning tool. & 4.67 & Strongly Agree\\
3 & The features of StatiCalcs are beneficial for practice and problem solving. & 4.76 & Strongly Agree\\
4 & I trust the accuracy of the calculations generated by StatiCalcs. & 4.56 & Strongly Agree\\
5 & Compared to textbooks, modules, and online videos, StatiCalcs provides a more convenient learning experience. & 4.46 & Strongly Agree\\
\multicolumn{2}{l}{\textbf{Overall Weighted Mean}} & \textbf{4.63} & \textbf{Strongly Agree}\\
\hline
\end{tabular}
\end{table}

The findings reveal that the student respondents were highly satisfied with the developed StatiCalcs platform, obtaining an overall weighted mean of 4.63 interpreted as ``Strongly Agree.'' The statement ``The features of StatiCalcs are beneficial for practice and problem solving'' obtained the highest weighted mean of 4.76, indicating that the respondents recognized the usefulness of the platform in reinforcing engineering concepts and solving Statics problems. Meanwhile, the statement comparing StatiCalcs with traditional learning resources obtained the lowest weighted mean of 4.46, although it still remained within the ``Strongly Agree'' interpretation range.

These findings align with \cite{johnson_customer_2001}'s User Satisfaction Model, which explains that users develop positive perceptions toward systems that effectively support their needs, expectations, and learning tasks. Similarly, \cite{pedraza_assessing_2025} emphasized that digital learning tools improve learner satisfaction by providing flexible access, interactive learning experiences, and immediate support for practice and engagement.

%==========Instructors’ Perception on Usability=========================

\section{Level of Perception of Instructors Regarding StatiCalcs in Terms of Usability}

\begin{table}[H]
\centering
\caption{Weighted Mean of Instructors’ Perception on Usability}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{cp{6.5cm}cc}
\hline
\textbf{No.} & \textbf{Statement} & \textbf{Weighted Mean} & \textbf{Interpretation}\\
\hline
1 & The StatiCalcs website is easy to navigate. & 4.73 & Strongly Agree\\
2 & I can complete calculations quickly using the website. & 4.64 & Strongly Agree\\
3 & The website provides clear feedback after performing calculations. & 4.00 & Agree\\
4 & I can easily correct mistakes when entering inputs in the calculators. & 4.45 & Strongly Agree\\
5 & Overall, StatiCalcs is user-friendly. & 4.45 & Strongly Agree\\
\multicolumn{2}{l}{\textbf{Overall Weighted Mean}} & \textbf{4.45} & \textbf{Strongly Agree}\\
\hline
\end{tabular}
\end{table}


The results indicate that the instructor respondents strongly agreed that StatiCalcs is usable and easy to navigate, with an overall weighted mean of 4.45 interpreted as ``Strongly Agree.''. The statement ``The StatiCalcs website is easy to navigate'' obtained the highest weighted mean of 4.73, indicating a highly favorable evaluation of the platform’s interface and usability. Meanwhile, the statement ``The website provides clear feedback after performing calculations'' obtained the lowest weighted mean of 4.00, interpreted as ``Agree.'' Despite this, the overall results demonstrate that instructors positively perceived the usability of the developed platform.

The findings support \cite{nielsen_10_1994}’s Usability Heuristics, which emphasize the importance of intuitive navigation, consistency, and efficiency in system design. Likewise, \cite{hassan_usability_2022} noted that user-friendly interfaces and simplified digital environments improve usability and allow users to interact with educational tools more effectively.

%==========Instructors’ Perception on Accessibility=========================

\section{Level of Perception of Instructors Regarding StatiCalcs in Terms of Accessibility}

\begin{table}[H]
\centering
\caption{Weighted Mean of Instructors’ Perception on Accessibility}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{cp{6.5cm}cc}
\hline
\textbf{No.} & \textbf{Statement} & \textbf{Weighted Mean} & \textbf{Interpretation}\\
\hline
1 & The text and visual elements of StatiCalcs are easy to read. & 4.55 & Strongly Agree\\
2 & The website works properly on different devices such as laptops and mobile phones. & 4.55 & Strongly Agree\\
3 & The layout makes important information easy to locate. & 4.18 & Agree\\
4 & The explanations and results provided by the website are easy to understand. & 4.45 & Strongly Agree\\
5 & Overall, StatiCalcs is accessible under different conditions. & 3.73 & Agree\\
\multicolumn{2}{l}{\textbf{Overall Weighted Mean}} & \textbf{4.29} & \textbf{Strongly Agree}\\
\hline
\end{tabular}
\end{table}

The results indicate that the instructor respondents positively evaluated the accessibility of StatiCalcs, obtaining an overall weighted mean of 4.29 interpreted as ``Strongly Agree.'' The statements ``The text and visual elements of StatiCalcs are easy to read'' and ``The website works properly on different devices such as laptops and mobile phones'' both obtained the highest weighted mean of 4.55, interpreted as ``Strongly Agree.'' Meanwhile, the statement ``Overall, StatiCalcs is accessible under different conditions'' obtained the lowest weighted mean of 3.73, interpreted as ``Agree.''

These findings are consistent with the Web Content Accessibility Guidelines (WCAG 2.1) of the \cite{world_wide_web_consortium_w3c_web_nodate}, which emphasize readability, device compatibility, and accessibility across varying user conditions. Furthermore, \cite{abumandour_applying_2021} highlighted that web-based learning environments should promote flexibility and accessibility to support effective learning experiences in engineering education.

%===============================

\section{Level of Perception of Instructors Regarding StatiCalcs in Terms of Satisfaction}


\begin{table}[H]
\centering
\caption{Weighted Mean of Instructors’ Perception on Satisfaction}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{cp{6.5cm}cc}
\hline
\textbf{No.} & \textbf{Statement} & \textbf{Weighted Mean} & \textbf{Interpretation}\\
\hline
1 & I am satisfied with my overall experience using StatiCalcs. & 4.27 & Strongly Agree\\
2 & StatiCalcs meets my expectations as a supplementary learning tool. & 4.27 & Strongly Agree\\
3 & The features of StatiCalcs are beneficial for practice and problem solving. & 4.36 & Strongly Agree\\
4 & I trust the accuracy of the calculations generated by StatiCalcs. & 4.45 & Strongly Agree\\
5 & Compared to textbooks, modules, and online videos, StatiCalcs provides a more convenient learning experience. & 3.55 & Agree\\
\multicolumn{2}{l}{\textbf{Overall Weighted Mean}} & \textbf{4.18} & \textbf{Agree}\\
\hline
\end{tabular}
\end{table}

The results reveal that the instructor respondents positively perceived StatiCalcs in terms of satisfaction, obtaining an overall weighted mean of 4.18 interpreted as ``Agree.'' The statement ``I trust the accuracy of the calculations generated by StatiCalcs'' obtained the highest weighted mean of 4.45, interpreted as ``Strongly Agree,'' indicating strong confidence in the reliability and correctness of the platform’s computational features. Meanwhile, the statement ``Compared to textbooks, modules, and online videos, StatiCalcs provides a more convenient learning experience'' obtained the lowest weighted mean of 3.55, interpreted as ``Agree.''

These findings support \cite{johnson_customer_2001}’s User Satisfaction Model, which explains that users develop satisfaction when a system effectively meets their expectations and supports their learning needs. Similarly, \cite{gfrerer_teaching_2023} emphasized that digital engineering tools and automated learning systems enhance learning experiences by improving analytical understanding and supporting more efficient problem-solving processes.
%===============================Preference of Students
\section{The Preference of Students for Different Learning Resources in Studying Statics of Rigid Bodies}

\begin{table}[H]
\centering
\caption{Weighted Mean and Ranking of Students’ Preference for Learning Resources}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c p{5.5cm} c c}
\hline
\textbf{Rank} & \textbf{Learning Resource} & \textbf{Weighted Mean} & \textbf{Interpretation} \\
\hline
1 & StatiCalcs & 4.55 & Most Preferred \\
2 & Textbooks & 4.52 & Most Preferred \\
3 & YouTube Videos & 4.33 & Most Preferred \\
4 & Modules & 4.32 & Most Preferred \\
5 & Online Engineering Tools & 4.00 & Preferred \\
6 & Lecture Notes & 3.98 & Preferred \\
\hline
\end{tabular}
\end{table}

The findings reveal that StatiCalcs obtained the highest weighted mean of 4.55, interpreted as ``Most Preferred,'' indicating that the student respondents highly favored the developed web-based learning tool as a supplementary resource in studying Statics of Rigid Bodies. This suggests that students prefer interactive and technology-supported learning environments that provide accessibility, visualization, and guided problem-solving features. This finding supports the studies of \cite{szymkowiak_information_2021} and \cite{tawafak_innovative_2021}, who emphasized that Generation Z learners respond positively to technology-based and interactive learning platforms that enhance engagement and comprehension.

Textbooks ranked second with a weighted mean of 4.52, followed by YouTube Videos and Modules with weighted means of 4.33 and 4.32, respectively, all interpreted as ``Most Preferred.'' Meanwhile, Online Engineering Tools and Lecture Notes obtained the lowest weighted means of 4.00 and 3.98, interpreted as ``Preferred.'' The findings indicate that although traditional learning resources remain valuable, students tend to favor more interactive and digitally accessible learning tools. This aligns with the findings of \cite{pedraza_assessing_2025}, which stated that digital learning platforms improve student engagement by providing flexible access, interactive content, and immediate learning support.

%===============================Preference of instructors


\section{The Preference of Instructors for Different Learning Resources in Studying Statics of Rigid Bodies}

\begin{table}[H]
\centering
\caption{Weighted Mean and Ranking of Instructors’ Preference for Learning Resources}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{c p{6cm} c c}
\hline
\textbf{Rank} & \textbf{Learning Resource} & \textbf{Weighted Mean} & \textbf{Interpretation} \\
\hline
1 & Textbooks (e.g., Hibbeler, Beer and Johnston) & 4.45 & Most Preferred \\
2 & Lecture Notes (e.g., class notes, handouts) & 3.82 & Preferred \\
3 & StatiCalcs (web-based learning tool) & 3.73 & Preferred \\
4 & Modules (e.g., instructor-provided materials) & 3.64 & Preferred \\
4 & Online Engineering Tools (e.g., SkyCiv, Mathalino) & 3.64 & Preferred \\
5 & YouTube Videos (e.g., educational tutorials) & 3.18 & Neutral \\
\hline
\end{tabular}
\end{table}

The results indicate that textbooks obtained the highest weighted mean of 4.45, interpreted as ``Most Preferred,'' showing that instructor respondents still highly value traditional academic resources in teaching Statics of Rigid Bodies. Lecture Notes ranked second with a weighted mean of 3.82, while StatiCalcs ranked third with a weighted mean of 3.73, both interpreted as ``Preferred.'' The findings suggest that instructors continue to rely on conventional instructional materials while also recognizing the usefulness of technology-supported learning tools such as StatiCalcs. This supports the findings of Abumandour, who emphasized that engineering education increasingly integrates traditional instruction with web-based and technology-assisted learning systems to enhance flexibility and accessibility.

Meanwhile, Modules and Online Engineering Tools both obtained weighted means of 3.64, interpreted as ``Preferred,'' while YouTube Videos obtained the lowest weighted mean of 3.18, interpreted as ``Neutral.'' The findings imply that instructors prefer structured and academically oriented learning materials over less formal online resources. Nevertheless, the positive evaluation of StatiCalcs indicates acceptance of web-based educational tools that support engineering instruction and problem-solving. This aligns with the findings of \cite{gfrerer_teaching_2023} and \cite{sollradl_symbolic_nodate}, who highlighted that digital engineering tools and automated learning systems can improve analytical understanding and support more effective learning experiences in Engineering Mechanics.


%TTEST==================

\section{Significant Difference Between the Perceptions of Students and Instructors Regarding StatiCalcs}

\begin{table}[H]
\centering
\caption{Test of Significant Difference Between the Perceptions of Students and Instructors}
\small
\renewcommand{\arraystretch}{1.3}

\begin{tabular}{lccccc}
\hline
\textbf{Variable} &
\textbf{\makecell{Student\\Mean}} &
\textbf{\makecell{Instructor\\Mean}} &
\textbf{P-value} &
\textbf{Decision} &
\textbf{Interpretation}\\
\hline
Usability & 4.66 & 4.45 & 0.19 & Accept Ho & No Significant Difference\\
Accessibility & 4.65 & 4.29 & 0.22 & Accept Ho & No Significant Difference\\
Satisfaction & 4.63 & 4.18 & 0.06 & Accept Ho & No Significant Difference\\
\hline
\end{tabular}
\end{table}

The student respondents obtained weighted mean values of 4.66 for usability, 4.65 for accessibility, and 4.63 for satisfaction. These results indicate that the students positively perceived \textit{StatiCalcs} as a usable, accessible, and satisfactory web-based learning tool for studying Statics of Rigid Bodies. Among the evaluated variables, usability obtained the highest weighted mean, suggesting that the respondents found the platform easy to navigate and user-friendly.

On the other hand, the instructor respondents obtained weighted mean values of 4.45 for usability, 4.29 for accessibility, and 4.18 for satisfaction. The results likewise reflect positive perceptions toward the developed learning tool. Furthermore, all computed p-values were greater than the 0.05 level of significance; therefore, the null hypothesis was accepted. This indicates that there was no significant difference between the perceptions of students and instructors regarding the usability, accessibility, and satisfaction of \textit{StatiCalcs}.

%======== comments

\section{Respondents’ Comments and Suggestions Regarding StatiCalcs}

The respondents provided several comments, suggestions, and observations regarding the developed \textit{StatiCalcs} platform. Most responses expressed positive perceptions toward the usability, accessibility, and usefulness of the website as a supplementary learning tool for studying Statics of Rigid Bodies. Several respondents described the platform as user-friendly, easy to navigate, fast, convenient, and effective for checking answers, reviewing lessons, and understanding engineering concepts. Some respondents also expressed appreciation for the integration of calculators, modules, formulas, and learning materials within a single platform. Other responses such as ``Hawud kaayo,'' ``Kamo palang jud,'' ``Bagsik,'' and ``Ayos'' further reflected the respondents’ positive reception and satisfaction toward the platform.

Many respondents also emphasized that the platform would be highly beneficial for engineering students, especially those struggling to understand Statics and related engineering subjects. Some respondents mentioned that they wished the website had already existed during their earlier years in engineering because of its usefulness and convenience in reviewing lessons and solving problems.

Despite the positive feedback, respondents also provided several recommendations for future improvements of the platform. Common suggestions included the addition of step-by-step solutions, more examples and practice problems, formula displays, diagrams, and video tutorials to improve learning support and usability. Other respondents recommended improving navigation features by adding back and next buttons, enhancing formula visibility, implementing input validation, strengthening mobile accessibility, and supporting additional units of measurement. Several respondents also suggested expanding the scope of the platform by incorporating additional engineering topics such as Dynamics of Rigid Bodies and Mechanics of Deformable Bodies, as well as improving visualization features for better conceptual understanding.

A few respondents also identified certain limitations and technical concerns within the platform. These included incomplete content in some modules, assumptions and restrictions in certain calculators, and issues involving the concurrent force calculator when solving for unknown angles. Some respondents also noted that the platform is more appropriate as a supplementary reference tool rather than a replacement for fundamental learning and conceptual understanding. Nevertheless, the overall responses indicate that the respondents positively received the platform and recognized its strong potential for further enhancement and broader educational application.

\section{User Testing Period Overview}

Over the 15-day testing period from May 1 to May 15, 2026, all system
activity was monitored and recorded on a daily basis. The system
processed a total of \textbf{16,756 requests} with a \textbf{0\% error
rate}, reflecting consistent and reliable performance throughout.

\vspace{1em}

\begin{table}[H]
\centering
\caption{Summary of Key Metrics}
\begin{tabular}{|c|c|}
\hline
\textbf{Metric} & \textbf{Value} \\
\hline
Testing Period & May 1 -- May 15, 2026 \\
\hline
Total Requests & \textbf{16,756} \\
\hline
Total Errors & 0 (0\%) \\
\hline
Peak Single-Day Requests & $\sim$2,740 --- May 11, 7:00 PM \\
\hline
Average Daily Requests & $\sim$1,197 \\
\hline
Blocked Requests & Not Available (Log Drains not connected) \\
\hline
\end{tabular}
\end{table}


\begin{figure}[H]
\centering
\begin{tikzpicture}
\begin{axis}[
    ybar,
    bar width=0.4cm,
    width=\textwidth,
    height=8cm,
    xlabel={Date (May)},
    ylabel={Total Requests},
    xlabel style={font=\small\bfseries, yshift=-4pt},
    ylabel style={font=\small\bfseries, xshift=-4pt},
    symbolic x coords={1,2,3,4,5,6,7,8,9,10,11,12,13,14,15},
    xtick=data,
    x tick label style={font=\footnotesize},
    y tick label style={font=\footnotesize},
    ymin=0,
    ymax=3200,
    ytick={0, 500, 1000, 1500, 2000, 2500, 3000},
    ymajorgrids=true,
    grid style={dashed, gray!30},
    axis line style={gray!50},
    nodes near coords,
    nodes near coords align={vertical},
    every node near coord/.append style={font=\footnotesize\bfseries, color=black},
    enlarge x limits=0.04,
]
\addplot[
    fill=blue!50!black!70,
    draw=blue!50!black!90,
    fill opacity=0.85
] coordinates {
    (1,   20)
    (2,  180)
    (3,   90)
    (4,   80)
    (5,  130)
    (6,   60)
    (7,  510)
    (8,  200)
    (9,  1000)
    (10, 200)
    (11, 2700)
    (12, 150)
    (13, 100)
    (14, 400)
    (15, 250)
};
\end{axis}
\end{tikzpicture}
\caption{Daily Request Volume --- May 1--15, 2026}
\label{fig:daily_requests}
\end{figure}


Over the 15-day testing period, the system processed a total of 16,756 requests with a 0\% error rate. A request refers to any single interaction sent to the system, such as a user submitting a form, loading a page, or triggering a feature, that the system must receive and respond to.

Usage was not uniform throughout the period. Most days recorded low activity ranging between 60 and 200 requests, averaging around 1,197 requests per day. May 9 stood out with 1,000 requests, roughly five times the typical daily volume, while May 11 represented the clear peak of the entire period at 2,700 requests. Despite this surge, the system absorbed the load without any slowdowns or reported failures, proving it can handle sudden heavy usage reliably.

Activity dropped sharply on May 12 and 13 to just 150 and 100 requests respectively, consistent with a weekend pattern and suggesting the traffic was driven by real users rather than automated processes. Usage recovered toward the end of the period, with May 14 recording 400 requests and May 15 closing at 250.

Overall, the system demonstrated strong stability and reliability throughout the entire testing window.



%